---
title: 'ERHS 535 Homework #2'
author: "Sarah Derda"
date: "Due Date: October 2, 2020"
output: word_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Section 1: Description of Data

This data is describing various airline companies safety records and whether or 
not these airlines have experienced incidents, fatal incidents, and/or fatalities
based on data from the years 1989 to 1999 and 2000 to 2014. The airline safety data
was collected from the "Aviation Safety Network's Database." The coder named
Nate Silver broke the data into the two previosuly mentioned time periods. Silver
wanted to look into whether people were less inclined to fly a particular airline 
because of their crash history - whether it was their fault or not. He had also
accounted for the number of available seat kilometers since one airline could fly
more or less than another ultimately adjusting this value. He also looked into the 
correlation between the time periods to see if there was a huge difference.
He found there was no correlation. Along with Silver's findings, I think it would 
be useful to look into a portion of which airlines had the least and the most incidents, fatal 
incidents, and/or fatalities overall. It would be interesting to look into the 
combined occurrences as well. Some researchers might be interested in this data to
see if there patterns across other time periods. If the data would be available, I
think it would be interesting to look into the impact of a pandemic, such as the
occurrence of COVID-19, and see if it affected airline safety based on fatalities, 
fatal incidents, and/or incidents. For this assignment, I'll be focusing on only 
non-U.S. airlines. The asterisk for each airline means that regional subsidiaries were
included. As for the format of the data, it is a comma-separated value file, which 
is a delimited text file that uses a comma delimiter to separate values or fields 
within the dataset.


## Section 2: Reading the Data into RStudio

In order to read in this dataset, you would need to install the "readr" package
and call upon it's library. There are multiple ways you could import the "csv" data
into RStudio. One way is by saving the file onto your computer to a specific name
and importing it using the "Import Dataset" button under "File." A second way is
calling upon the directory where it is located. Another way is by using the 
"paste0" function to create an object with the url since we're importing it through 
GitHub and then reading it into RStudio with "read_csv" by inputting the file name
used to import the data from the internet. The last option is shown below.

I did not need to use any special options to upload this data. The dates from the
original dataset were sectioned into two time periods mentioned previously, so no
dates needed to be formatted as they were already included into the variable name. 
I did end up renaming some of the columns for easier readability. In order to look 
at non-U.S. airlines only, I filtered out all non-U.S. airlines. This was a fairly
simple dataset, so I didn't need to arrange anything using arrange() or add any 
other variables using mutate() for my interests.


```{r libraries, include = TRUE, results = FALSE}

# Upload libraries to use functions.

library(knitr)
library(readr)
library(dplyr)
library(tidyverse)
library(ggplot2)

```

```{r airline dataset}

airline_url <- paste0("https://raw.githubusercontent.com/fivethirtyeight/data/master/airline-safety/airline-safety.csv")

airline <- read_csv(file = airline_url)

head(airline)

summary(airline)

airline2 <- airline %>% 
  rename(Airline = airline,
         PAX_Carrying_Capacity = avail_seat_km_per_week,
         Incidents_85_99 = incidents_85_99,
         Fatal_Accidents_85_99 = fatal_accidents_85_99,
         Fatalities_85_99 = fatalities_85_99,
         Incidents_00_14 = incidents_00_14,
         Fatal_Accidents_00_14 = fatal_accidents_00_14,
         Fatalities_00_14 = fatalities_00_14) %>%
  filter(Airline %in% c("Aer Lingus", "Aeroflot*", "Aerolineas Argentinas", "Aeromexico*",
                        "Air Canada", "Air France", "Air India*", "Air New Zealand*",
                        "Alitalia", "All Nippon Airways", "Austrian Airlines", "Avianca",
                        "British Airways", "Cathay Pacific", "China Airlines", "Condor",
                        "COPA", "Egyptair", "El Al", "Ethiopian Airlines", "Finnair",
                        "Garuda Indonesia", "Gulf Air", "Iberia", "Japan Airlines",
                        "Kenya Airways", "KLM*", "Korean Air", "LAN Airlines", "Lufthansa*",
                        "Malaysia Airlines", "Pakistan International", "Philippine Airlines",
                        "Qantas*", "Royal Air Maroc", "SAS*", "Saudi Arabian", "Singapore Airlines",
                        "South African", "Sri Lankan / AirLanka", "SWISS*", "TACA", "TAM",
                        "TAP - Air Portugal", "Thai Airways", "Turkish Airlines", "Vietnam Airlines",
                        "Virgin Atlantic", "Xiamen Airlines"))

```


## Section 3: Characteristics of the Data

There are `r nrow(airline2)` rows and `r ncol(airline2)` columns in the "airline2" dataset. 

```{r table}

airline_table <- data.frame(letters = c("Airline", "PAX_Carrying_Capacity", "Incidents_85_99",
                                   "Fatal_Accidents_85_99", "Fatalities_85_99", "Incidents_00_14",
                                   "Fatal_Accidents_00_14", "Fatalities_00_14"),
                       letters = c("Airline name", "Available seats per kilometer per week", "Incidents from 1985 to 1999", "Fatal accidents from 1985 to 1999", "Fatalities from 1985 to 1999", "Incidents from 2000 to 2014", "Fatal accidents from 2000 to 2014", "Fatalities from 2000 to 2014"),
                       letters = c("None", "None", "None", "None", "None", "None", "None", "None"))

kable(airline_table, align = c("l", "l", "l"),
      caption = "Airline Characteristics", 
      col.names = c("Column Name", "Description", "Units"))

```


## Section 4: Summary Statistics

There were no missing values from any of the airlines. The summaries are shown below.

```{r summarize}

sum(is.na(airline2))

airline_summaries <- airline2 %>%
  summarize(min_inc_00_14 = min(Incidents_00_14), max_inc_00_14 = max(Incidents_00_14),
            mean_inc_00_14 = mean(Incidents_00_14), na_inc_00_14 = sum(is.na(Incidents_00_14)),
            min_facc_00_14 = min(Fatal_Accidents_00_14), max_facc_00_14 = max(Fatal_Accidents_00_14),
            mean_facc_00_14 = mean(Fatal_Accidents_00_14), na_facc_00_14 = sum(is.na(Fatal_Accidents_00_14)),
            min_fat_00_14 = min(Fatalities_00_14), max_fat_00_14 = max(Fatalities_00_14),
            mean_fat_00_14 = mean(Fatalities_00_14), na_fat_00_14 = sum(is.na(Fatalities_00_14))) %>%
  pivot_longer(col = min_inc_00_14:na_fat_00_14,
             names_to = "Summary",
             values_to = "Value")

print(airline_summaries)

```


## Section 5: Visualizing Data

While first visualizing the data, I wanted to look into comparing incidents between 
the two time periods among non-U.S. airlines. This would allow us to get a better 
glimpse into the overall distrubution or frequency of incidents in non-U.S. airlines 
using histograms that are weighted by passenger carrying capacity (seats per kilometers 
per week). Secondly, by producing scatterplots showing which non-U.S. airlines had 
the most incidents, we have some insight into which *might* have more of a risk of an
incident occurring. Again, this is only examining incidents as that has some indication 
of something going wrong with the flight rather than looking into fatalities, which 
indicates that the incident or accident was fatal. Furthermore, these graphs 
could aid in looking at histories of an individual airline - U.S. or non-U.S based. 
Also, I think it would be interesting comparing two countries' airlines if a researcher 
wanted to examine countries that seem to portray more of a risk in safety. In producing 
these graphs, two good graphing techniques that were met were 1) high data density 
and 2) making the order meaningful since there were two time periods involved. High 
data density was met in the ink to data ratio within the graphs. The order was meaningful 
due to the fact that we're comparing the time period from 1985-1999 and 2000-2014.
However, it is important to note that when comparing these, the values along the x-axis
are different between the two time periods.

As we can see from the graphs, it looks like there were more incidents among all airlines
from 1985 to 1999 compared to the time period from 2000 to 2014. From 1985 to 1999,
Aeroflot had the highest number of incidents. From 2000 to 2014, Saudi Arabian had
the highest amount of incidents, but it was a far less number than Aeroflot.


```{r plots, fig.width = 8, fig.height = 6}

# Statistical Graphing - Histograms

p1 <- ggplot(airline2, aes(x = Incidents_85_99, y = ..density.., weight = PAX_Carrying_Capacity)) + 
          geom_histogram(binwidth = 1, color = "black", fill = "white") + 
          labs(x = "Incidents, 1985-1999", y = "Density", title = "Incidents per Trillion Available Seat Kilometers for All Non-U.S. Airlines, 1985-1999") + 
          theme_classic() + geom_density(alpha=.5, fill="pink")
p1

p2 <- ggplot(airline2, aes(x = Incidents_00_14, y = ..density.., weight = PAX_Carrying_Capacity)) + 
          geom_histogram(binwidth = 1, color = "black", fill = "white") + 
          labs(x = "Incidents, 2000-2014", y = "Density", title = "Incidents per Trillion Available Seat Kilometers for All Non-U.S. Airlines, 2000-2014") +
          theme_classic() + geom_density(alpha=.5, fill="pink")
p2

# Non-statistical Graphing - Scatterplots

p3 <- ggplot(airline2, aes(x = Incidents_85_99, y = Airline, size = PAX_Carrying_Capacity)) +
          geom_point(color = "red") + 
          labs(x = "Incidents, 1985-1999", y = "Airline", title = "Incidents per Trillion Available Seat Kilometers per Non-U.S. Airlines, 1985-1999") + 
          theme_classic() 
p3

p4 <- ggplot(airline2, aes(x = Incidents_00_14, y = Airline, size = PAX_Carrying_Capacity)) +
          geom_point(color = "red") + 
          labs(x = "Incidents, 2000-2014", y = "Airline", title = "Incidents per Trillion Available Seat Kilometers per Non-U.S. Airline, 2000-2014") + 
          theme_classic()
p4

```



