---
title: 'Homework #3'
author: "Sarah Derda"
date: "Due Date: October 16, 2020"
output: word_document
---

## Question #1: Setting Global Options  



```{r global_options&libraries}

knitr::opts_chunk$set(echo = TRUE, 
                      warning=FALSE, 
                      message=FALSE, 
                      error=FALSE)

library(readr)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(titanic)
library(stringr)

```

## Question #2: Editing HW#2 Plot  



```{r homework#2plots}

# Importing Data

airline_url <- paste0("https://raw.githubusercontent.com/fivethirtyeight/data/master/airline-safety/airline-safety.csv")
airline <- read_csv(file = airline_url)
head(airline)
summary(airline)

# Cleaning Data

airline2 <- airline %>% 
  rename(Airline = airline,
         PAX_Carrying_Capacity = avail_seat_km_per_week,
         Incidents_85_99 = incidents_85_99,
         Fatal_Accidents_85_99 = fatal_accidents_85_99,
         Fatalities_85_99 = fatalities_85_99,
         Incidents_00_14 = incidents_00_14,
         Fatal_Accidents_00_14 = fatal_accidents_00_14,
         Fatalities_00_14 = fatalities_00_14) %>%
  filter(Airline %in% c("Aer Lingus", "Aeroflot*", "Aerolineas Argentinas", "Aeromexico*",
                        "Air Canada", "Air France", "Air India*", "Air New Zealand*",
                        "Alitalia", "All Nippon Airways", "Austrian Airlines", "Avianca",
                        "British Airways", "Cathay Pacific", "China Airlines", "Condor",
                        "COPA", "Egyptair", "El Al", "Ethiopian Airlines", "Finnair",
                        "Garuda Indonesia", "Gulf Air", "Iberia", "Japan Airlines",
                        "Kenya Airways", "KLM*", "Korean Air", "LAN Airlines", "Lufthansa*",
                        "Malaysia Airlines", "Pakistan International", "Philippine Airlines",
                        "Qantas*", "Royal Air Maroc", "SAS*", "Saudi Arabian", "Singapore Airlines",
                        "South African", "Sri Lankan / AirLanka", "SWISS*", "TACA", "TAM",
                        "TAP - Air Portugal", "Thai Airways", "Turkish Airlines", "Vietnam Airlines",
                        "Virgin Atlantic", "Xiamen Airlines"))

```


```{r plots, fig.width = 10, fig.height = 6}

# Plotting Data

ggplot(airline2, aes(x = Incidents_00_14, y = Airline, size = PAX_Carrying_Capacity)) +
          geom_point(color = "red") + 
          labs(x = "Incidents, 2000-2014", y = "Airline", title = "Incidents per Trillion Available Seat Kilometers per Non-U.S. Airline, 2000-2014") + 
          theme_classic()

# Editing Plot

ggplot(airline2, aes(x = Incidents_00_14, y = Airline, 
                           color = PAX_Carrying_Capacity,
                           size = PAX_Carrying_Capacity)) +
          geom_point() + 
          labs(x = "Incidents, 2000-2014", y = "Non-U.S. Airline", title = "Incidents per Trillion Available Seat Kilometers per Non-U.S. Airline, 2000-2014", size = "Passenger Carrying Capacity") + 
          theme_minimal() +
          theme(axis.title = element_text(size = 10), 
                plot.title = element_text(size = 10), 
                axis.text.y = element_text(size  = 5),
                legend.title = element_text(size = 10),
                legend.text = element_text(size = 5))

```


Among the six guidelines for good graphics, the first graph above depicts high data density 
and using clear, meaningful labels. Aiming for high data density allows the viewer
to easily see the message you're trying to convey as a coder. The higher the data
to ink ratio is, the less "ink" there will be to process, which takes less time to
read overall. Using clear, meaninful labels is also beneficial for readability.
Providing descriptive labels for viewers allow them to interpret the graph better
by clearly stating what each aspect of the graph conveys.

The second graph was edited. There was a background grid added in order to look at
the points more easily between the airlines and how many incidents they had per 
passenger carrying capacity. The text was also enlarged to make it easier to read. 
Furthermore, the color of the points were changed in order to depict the size of passenger carrying
capacity. Font sizes were also adjusted in order to distuinguish the titles and the
general text.

## Question #3: Titanic Dataset  



```{r titanic}

# Import Data & Observe

tt <- titanic_train
head(tt)

survdata <- tt %>%
  as.data.frame() %>%
  select("Survived", "Age")

class(survdata$Survived)

# Change Variable from Integer to Factor

survdata$Survived <- as.factor(survdata$Survived)
class(survdata$Survived)

levels(survdata$Survived)

# Relabel

survdata <- mutate(.data = survdata,
              Survived = fct_recode(.f = Survived,
                                  Died = "0",
                                  Survived = "1"))

# Top Ages of Survived and Died

top_surv <- survdata %>%
  filter(Survived == "Survived") %>%
  arrange(desc(Age)) %>%
  slice(1:10)

top_died <- survdata %>%
  filter(Survived == "Died") %>%
  arrange(desc(Age)) %>%
  slice(1:10)

# Merge to Form 10 Rows

join_surv <- full_join(top_surv, top_died)
print(join_surv)

# Plot

ggplot(survdata, aes(x = Age)) + 
        geom_histogram(fill = "white", color = "black") + 
        facet_wrap(~ Survived, ncol = 1) + 
        theme_few() +
        labs(y = "Frequency", 
             title = "Frequency Distribution of Age Among Deaths and Survivals")

# Two-Row Table Details

survdata_df <- survdata %>%
  group_by(Survived) %>%
  summarize(N = n(), mean = mean(Age, na.rm = TRUE), na = sum(is.na(Age)))
print(survdata_df)

# Stat Test

t.test(survdata$Age[survdata$Survived=="Died"], survdata$Age[survdata$Survived=="Survived"])
  
```

## Question #4: Word Puzzle  



```{r wordphrase}

words_alpha_url <- read_tsv("https://raw.githubusercontent.com/dwyl/english-words/master/words_alpha.txt", col_names = "word")

words_alpha <- words_alpha_url

wordphrase <- words_alpha %>%
  filter(str_detect(word, "^.{8}$")) %>%
  filter(str_detect(word, "^[Mm]")) %>%
  mutate(first = str_sub(word, 1, 4),
         second = paste0(str_sub(word, 6, 8), str_sub(word, 5, 5))) %>%
  filter(first %in% words_alpha$word & second %in% words_alpha$word)

```

I was not able to find the wordphrase. With the help of Nick, I was able to reduce
the list to 215 words; however, all of them look like they have a different sound
after moving the first letter of the second word to the end of the eight word phrase.
